---
title: "Replication_Assignment_inovick"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Instructions##
#You do not need to replicate ALL of the analyses presented in the paper, but at minimum you must replicate at least 3 analyses, including at least one descriptive statistical analysis and one inferential statistical analysis. As part of this assignment, you must also replicate to the best of your abilities at least one figure.

#Descriptive statistical analysis: All host preferences of mites from y-maze experiments. These are essentially the proportion of y-mazes in which the group A fly was infected.

#Inferential statistical analysis: Sex differences. To test which fly sex has a stronger response to parasites we measured the rate of CO2 production, a proxy for metabolic rate, of male and female flies either unexposed or exposed to mites using flow-through respirometry. Model comparison was carried out with the anova function (test=χ2) for lmer models and the Wald t-statistic for glmer models.

#Other analysis: Host reproductive status. A binomial test (binom.test) was used to test if mites 
disproportionately infected mated or virgin females.

#Figure: Fig. 2. Fly responses to mite exposure. A) Metabolic rates of male and female flies at rest or exposed to mites. B) Activity of male and female flies at rest or exposed to mites. One male fly was removed due to high activity. C) Metabolic rates of virgin and mated females at rest or exposed to mites. D) Activity of virgin and mated females at rest or exposed to mites. Boxplots represent minimum, 25th percentile, median, 75th percentile and maximum,.

#Include short description of the study and of the specific data and replication analyses you will be performing, to orient your reader. Outline (briefly) the goal of the original paper, the data set used, and the analyses conducted, then describe which you will replicate. You should also demonstrate how you read your datafile into R, and show a few lines of raw data in your output (e.g., using head()).

##Step 1. Reading the data into R:
#The data was available on Dryad as an .xlsx with 4 distinct tabs, holding different sets of data. I converted each tab to a .csv format, so now I have 4 different .csv files. These are respirometry data for male and female flies at rest, respirometry data for males and females exposed to mites, respirometry data for mated and virgin females at rest, and respirometry data for mated and virgin females exposed to mites. I have assigned each of them a name.
```{r}
mf_rest <- read.csv("Respirometry_data_male_and_female_at_rest.csv", header = TRUE, stringsAsFactors = FALSE)
head(mf_rest)

mf_exposed <- read.csv("Respirometry_data_male_and_female_exposed_to_mite.csv", header = TRUE, stringsAsFactors = FALSE)
head(mf_exposed)

fem_rest <- read.csv("Respirometry_data_mated_and_virgin_females_rest.csv", header = TRUE, stringsAsFactors = FALSE)
head(fem_rest)

fem_exposed <- read.csv("Respirometry_data_mated_and_virgin_exposed.csv", header = TRUE, stringsAsFactors = FALSE)
head(fem_exposed)
```

##Step 2. Load packages (or install them if I don't have them yet):
```{r}
library(ggplot2)
library(stats)
library(lme4)
library(dplyr)
install.packages("plotrix")
library(plotrix)
```

##Step 3. Descriptive statistical analysis: Taking the average weight of exposed male and female flies. 
#From section 3.2.1 in results. Female flies in the exposed experiment on average had 24% higher 
masses than male flies, 2.48 ± 0.06 mg and 1.95 ± 0.03 mg respectively. 
```{r}
summary(mf_exposed) #Use summary to get an idea of the data

male_exposed <- mf_exposed %>% filter(Sex == "M")
male_exposed
summary(male_exposed)
mean(male_exposed$Mass)

std.error(male_exposed$Mass)
```
#Mean mass for exposed males is 1.945, and standard error is 0.03 mg. It worked! Yeehaw! Now for the females:

```{r}
fem_exposed <- mf_exposed %>% filter(Sex == "F")
fem_exposed
summary(fem_exposed)
mean(fem_exposed$Mass)

std.error(fem_exposed$Mass)
```
#Mean mass for exposed females is 2.48, and standard error is 0.06 mg. Awesome!

##Step 4. Inferential statistical analysis: Sex differences. 

"To test which fly sex has a stronger response to parasites we measured the rate of CO2 production, a proxy for metabolic rate, of male and female flies either unexposed or exposed to mites using flow-through respirometry. Model comparison was carried out with the anova function (test=χ2) for lmer models and the Wald t-statistic for glmer models.

The MR of unexposed flies was best described with a normal distribution upon visual inspection, thus the lmer function was used for modeling. Unexposed females (N = 36) had substantially higher CO2 production rates than males (N = 36), 0.063 ± 0.003 μL/min versus 0.053 ± 0.003 μL/min respectively (16.3% higher), and sex was a 
significant  predictor  of  MR  among  unexposed  flies  (χ2=6.34, P = 0.012) (Fig. 2a)

The MR of the flies exposed to mites was best described using a gamma distribution (glmer, family=”Gamma”). In the exposed treatment,  the  MR  of  male  (0.073 ± 0.008  μL/min)  and  female (0.073 ± 0.007 μL/min) flies were nearly identical, and sex was not a significant predictor of MR (Wald t = 0.24, P = 0.81)"

#look at module 18 for this
#Looking at the MR of unexposed flies (at rest) and comparing differences based on sex using lmer function:
```{r}

```

#Looking at MR of exposed flies and comparing differences based on sex using gmer function:
```{r}

```


#Comparing exposed and exposed males and females using ANOVA function (test=χ2) for lmer models and Wald t-statistic for glmer models:
```{r}

```

#Recreating this figure: (insert link to picture)
```{r}
fem_exposed_mr <- fem_exposed$CO2.prod.uL.min
fem_rest_mr <- fem_rest$CO2.prod.uL.min
male_exposed_mr <- male_exposed$CO2.prod.uL.min
male_rest_mr <- male_rest$CO2.prod.uL.min

boxplot(fem_exposed_mr, fem_rest_mr, male_exposed_mr, male_rest_mr)

# create a data frame
variety=rep(LETTERS[1:7])
treatment=rep(c("high","low"),each=20)
note=seq(1:280)+sample(1:150, 280, replace=T)
data=data.frame(variety, treatment ,  note)
 
# grouped boxplot
ggplot(data, aes(x=variety, y=note, fill=treatment)) + 
    geom_boxplot()

DF2 <- data.frame(
  x = c(c(A1, A2, A3), c(B1, B2, B3)),
  y = rep(c("A", "B"), each = 15),
  z = rep(rep(1:3, each=5), 2),
  stringsAsFactors = FALSE
)
str(DF2)
# 'data.frame': 30 obs. of  3 variables:
#  $ x: num  1 2 9 6 4 5 1 9 2 3 ...
#  $ y: chr  "A" "A" "A" "A" ...
#  $ z: int  1 1 1 1 1 2 2 2 2 2 ...

cols <- rainbow(3, s = 0.5)
boxplot(x ~ z + y, data = DF2,
        at = c(1:3, 5:7), col = cols,
        names = c("", "A", "", "", "B", ""), xaxs = FALSE)
legend("topleft", fill = cols, legend = c(1,2,3), horiz = T)
```





###Stuff that didn't work...

#Calculating average weight for male flies at rest:
```{r}
male_rest <- mf_rest %>% filter(Sex == "M")
male_rest
summary(male_rest)
```
#Mean mass for males at rest is 2.208, which is closer, but still not the same. I will now try combining all males, both exposed and at rest, to see if I can get the same mean from the study. Since they are in two different .csv files, I will try and combine them into one column using the merge function.

#The merge function did not work for me, so I instead created a new .csv file by copying and pasting values from excel to create a combined file of males at rest and exposed:
```{r}
male_combined <- read.csv("combined_mf.csv", header = TRUE, stringsAsFactors = FALSE)
summary(male_combined$Mass)
```
#Mean mass of all males is 2.093, not 2.35