---
title: "Replication_Assignment_inovick"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#####

#####

##Introduction
#   The study containing the analyses I am attempting to replicate is "Extending the ecology of fear: Parasite-mediated sexual selection drives host response to parasites" by Horn et al, 2020. This study investigates non-consumptive effects of parasites (using ectoparasite Macrocheles subbadius) on potential hosts (Drosophila nigrospiracula). The authors use metabolic rate (MR) of host flies as a proxy for perceived risk of mite infection, recording the rate of CO2 production. They test two mutually exclusive hypotheses: either uneven costs of infection result in adaptions for stronger metabolic responses to mite exposure, or uneven risks of infection result in adaptions for stronger metabolic responses to mite exposure. Male flies will have higher cost of being parasitized due to their lower body mass compared to female flies. Virgin female flies will have a higher cost of being parasitized compared to mated females. These differences also affect infection risk, as mites preferentially infect female flies over males, and mated females over virgin females.
  The dataset used in this study recorded data in the following categories: individual code, respirometry chamber, sex, mass, CO2 production in ml/min, VCOsd, age, activity in voltage, mite infection, and batch. Not all of these variables are investigated in my replication analysis. The tests performed in this study are age-matched and pair-wise choice tests, binomial tests, linear mixed effects models using the lmer function, generalized linear mixed effects models using the glmer function, and backwards model selection. Additionally, after performing an lmer test, the anova function was used to compare models, and the Wald t-statistic test was used for comparing glmer models as they are in a Gamma distribution. The study also uses a Shapiro test on the residuals of models for normality.
  
  The analyses that I will be replicating are as follows:
#Descriptive statistical analysis: Taking average weight of exposed male and female flies.

#Inferential statistical analysis: Sex differences. To test which fly sex has a stronger response to parasites we measured the rate of CO2 production, a proxy for metabolic rate, of male and female flies either unexposed or exposed to mites using flow-through respirometry. Model comparison was carried out with the anova function (test=χ2) for lmer models and the Wald t-statistic for glmer models. I did lmer tests on the unexposed group, and then ran anovas to compare those models. I only compared variables of unexposed flies to see which variables are good predictors.

#Other analysis: Mating status. MR response of mated and virgin females to mite risk. I will do lmer test on mated and virgin females.

#Other OTHER analysis: Shapiro-Wilk test of normality

#SUPER OTHER other analysis: AIC to compare prediction error between models

#Figure: Fig. 2. Fly responses to mite exposure. A) Metabolic rates of male and female flies at rest or exposed to mites. B) Activity of male and female flies at rest or exposed to mites. One male fly was removed due to high activity. C) Metabolic rates of virgin and mated females at rest or exposed to mites. D) Activity of virgin and mated females at rest or exposed to mites. Boxplots represent minimum, 25th percentile, median, 75th percentile and maximum,.


##Step 1. Reading the data into R:
#The data was available on Dryad as an .xlsx with 4 distinct tabs, holding different sets of data. I converted each tab to a .csv format, so now I have 4 different .csv files. Eventually I ended up splitting these further into more .csv files, containing each variable separately to help with analysis. These original files are respirometry data for male and female flies at rest, respirometry data for males and females exposed to mites, respirometry data for mated and virgin females at rest, and respirometry data for mated and virgin females exposed to mites. I have assigned each of them a name.

#Loading in respirometry data for male and female flies at rest:
```{r}
mf_rest <- read.csv("Respirometry_data_male_and_female_at_rest.csv", header = TRUE, stringsAsFactors = FALSE)
head(mf_rest)
```

#Loading in respirometry data for males and females exposed to mites:
```{r}
mf_exposed <- read.csv("Respirometry_data_male_and_female_exposed_to_mite.csv", header = TRUE, stringsAsFactors = FALSE)
head(mf_exposed)
```

#Loading in respirometry data for mated and virgin females at rest:
```{r}
fem_rest <- read.csv("Respirometry_data_mated_and_virgin_females_rest.csv", header = TRUE, stringsAsFactors = FALSE)
head(fem_rest)
```

#Loading in respirometry data for mated and virgin females exposed to mites:
```{r}
fem_exposed <- read.csv("Respirometry_data_mated_and_virgin_exposed.csv", header = TRUE, stringsAsFactors = FALSE)
head(fem_exposed)
```

##Step 2. Load packages (or install them if I don't have them yet):
```{r}
library(ggplot2)
library(stats)
library(lme4)
library(dplyr)
install.packages("plotrix")
library(plotrix)
install.packages("lmerTest")
library(lmerTest)
install.packages("aod")
library(aod)
```

##Step 3. Descriptive statistical analysis: Taking the average weight of exposed male and female flies. 
#From section 3.2.1 in results. Female flies in the exposed experiment on average had 24% higher 
masses than male flies, 2.48 ± 0.06 mg and 1.95 ± 0.03 mg respectively. 
```{r}
summary(mf_exposed) #Use summary to get an idea of the data

male_exposed <- mf_exposed %>% filter(Sex == "M")
male_exposed
summary(male_exposed)
mean(male_exposed$Mass)

std.error(male_exposed$Mass)
```
#Mean mass for exposed males is 1.945, and standard error is 0.03 mg. It worked! Yeehaw! Now for the females:

```{r}
fem_exposed <- mf_exposed %>% filter(Sex == "F")
fem_exposed
summary(fem_exposed)
mean(fem_exposed$Mass)

std.error(fem_exposed$Mass)
```
#Mean mass for exposed females is 2.48, and standard error is 0.06 mg. Awesome!

##Step 4. Inferential statistical analysis: Sex differences. 

"To test which fly sex has a stronger response to parasites we measured the rate of CO2 production, a proxy for metabolic rate, of male and female flies either unexposed or exposed to mites using flow-through respirometry. Model comparison was carried out with the anova function (test=χ2) for lmer models and the Wald t-statistic for glmer models.

The MR of the flies exposed to mites was best described using a gamma distribution (glmer, family=”Gamma”). In the exposed treatment,  the  MR  of  male  (0.073 ± 0.008  μL/min)  and  female (0.073 ± 0.007 μL/min) flies were nearly identical, and sex was not a significant predictor of MR (Wald t = 0.24, P = 0.81)"

###From Collin Horn: I use the lme4 package, and have started using lmertest instead of many anovas, it seems easier and less likely to make user errors (to retrieve P-values).
##My response: Using lmertest is giving me somewhat different results than what was published in the paper, since the paper was written using several anovas originally. So I am going to do a bunch of anovas to try and get the same answer.

So for the situation looking at fly sex without mites I made a model. Fly age is related to batch, btw.
```{r}
EEErestMF <- lmer(CO2.prod.uL.min ~ Age + Sex + Mass +(batch|Age), data = mf_rest)
EEErestMF
summary(EEErestMF)
anova(EEErestMF)
```
#"Mass  is  not  a  significant  predictor  of  MR (χ2<0.001, P = 0.99)": 
```{r}
lmernomass <- lmer(CO2.prod.uL.min ~ Sex + Act + chamber +(batch|Age), data = mf_rest)
lmer <- lmer(CO2.prod.uL.min ~ Mass + Sex + Act + chamber +(batch|Age), data = mf_rest)

anova(lmer, lmernomass)
```
#"Nor was mass a significant predictor in models without sex (χ2=0.43, P = 0.51)." This one includes mass as a predictor and excludes sex for both, excluding mass for one:
```{r}
restdocmass <-lmer(CO2.prod.uL.min ~ Mass + Act + chamber +(batch|Age), data = mf_rest)
summary(restdocmass)

restdocnomass <- lmer(CO2.prod.uL.min ~ Act + chamber +(batch| Age), data = mf_rest)
summary(restdocnomass)

anova(restdocmass, restdocnomass)
```

#Using sex as a predictor, excluding mass:
```{r}
restsex <- lmer(CO2.prod.uL.min ~ Sex + Act + chamber + (batch | Age), data = mf_rest)
summary(restsex)
```
#Using Akaike Information Criterion to estimate prediction error. This is to see if sex or mass are better predictors.
```{r}
AIC(restsex, restdocmass)
```
#From this, we can see that sex is a better predictor than mass. We do separate tests of mass and sex to avoid colinearity because of sexual dimorphism and the large difference in mass between the sexes.

#Sex is a significant predictor of MR among unexposed flies (at rest): "The MR of unexposed flies was best described with a normal distribution upon visual inspection, thus the lmer function was used for modeling. Unexposed females (N = 36) had substantially higher CO2 production rates than males (N = 36), 0.063 ± 0.003 μL/min versus 0.053 ± 0.003 μL/min respectively (16.3% higher), and sex was a significant  predictor  of  MR  among  unexposed  flies  (χ2=6.34, P = 0.012). Success!
```{r}
restsex <- lmer(CO2.prod.uL.min ~ Sex + Act + chamber + (batch | Age), data = mf_rest)
restnosex <- lmer(CO2.prod.uL.min ~ Act + chamber + (batch | Age), data = mf_rest)
summary(restnosex)

anova(restsex, restnosex)
```

#Comparing sex including and excluding activity, looking at activity as a predictor. Not included in study but trying to be thorough.
```{r}
sexnoact <- lmer(CO2.prod.uL.min ~ Sex + chamber + (batch | Age), data = mf_rest)
summary(sexnoact)
anova(restsex, sexnoact)

#This is new minimal model
```

##"Respirometry chamber was a significant predictor of MR (χ2=12.3, P = 0.0004)."
#Comparing sex including chamber and excluding chamber:
```{r}
sexnocham <- lmer(CO2.prod.uL.min ~ Sex + Act + (batch | Age), data = mf_rest)
summary(sexnocham)

anova(restsex, sexnocham)
```
#P-value is the same, and chisq value is very close (they got 12.3, I got 12.6). I got some kind of error that maximum number of function evaluations exceeded, so I have to think that has something to do with the discrepancy. I will try it with including mass as well:

```{r}
lmernocham <- lmer(CO2.prod.uL.min ~ Mass + Sex + Act + (batch| Age), data = mf_rest)

anova(lmer, lmernocham)
```
#Close but not exactly, probably because of an error saying the model is failing to converge with 7  negative eigenvalues. This gives a p-value of 0.0002 and a chisq value of 13.556, so I think it was originally done by excluding mass like my first attempt. 

#Comparing no activity in model using sex as predictor, and no activity and no sex:
```{r}
nosexact <- lmer(CO2.prod.uL.min ~ chamber + (batch | Age), data = mf_rest)
summary(nosexact)

anova(sexnoact, nosexact)
```
#Comparing no activity and no chamber to no activity with sex as a predictor
```{r}
sexnoactcham <- lmer(CO2.prod.uL.min ~ Sex + (batch | Age), data = mf_rest)

anova(sexnoact, sexnoactcham)
```

#Additional statistical test: Doing a Shapiro-Wilk normality test:
```{r}
shapiro.test(resid(sexnoact))
```
#Histogram of CO2 respirometry:
```{r}
CO2_Production <- mf_rest$CO2.prod.uL.min
hist(CO2_Production, plot = TRUE)
```





##With the exposed condition the data was slightly right skewed and we used a Gamma distribution and the glmer function. Trying to figure out how to get t statistic and fucking around with glmer
```{r}
EEEexpMF <- glmer(CO2.prod.uL.min ~ Age + Sex + Mass + (chamber|batch), family= "Gamma", data = mf_exposed)
EEEexpMF

glmer <- glmer(CO2.prod.uL.min ~ Sex + Act + Mass + (chamber|batch), family= "Gamma", data = mf_exposed)
glmer

summary(glmer)
glmernosex <- glmer(CO2.prod.uL.min ~ Act + Mass + (chamber|batch), family= "Gamma", data = mf_exposed)
glmernosex
summary(glmernosex)
anova(glmer, glmernosex, test = "Wald")
wald.test(glmer, glmernosex, Terms = mf_exposed$sex)
summary(glmernosex, glmer)

glmernomass <- glmer(CO2.prod.uL.min ~ Act + Sex + (chamber|batch), family= "Gamma", data = mf_exposed)
summary(glmernomass)
EEEexpMF<- lmer(CO2.prod.uL.min~+Sex+Mass+(chamber|batch), data=mf_exposed)
summary(EEEexpMF)

glmernosex<- lmer(CO2.prod.uL.min~Age+Mass+(chamber|batch), data=mf_exposed)
summary(glmernosex)

glmernomass <-lmer(CO2.prod.uL.min~Age+Sex+(chamber|batch), data=mf_exposed)
summary(glmernomass)

```
##Exposed treatment: mass specific analysis from Collin Horn's exposed analysis doc. fucking around:
```{r}
expmassglmer <- glmer(CO2.prod.uL.min ~ Mass + Act + chamber +(batch|Age), family= "Gamma", data = mf_exposed)
summary(expmassglmer)

ulminmg ~ Sex * Act + chamber + (batch | date)
Ulmin ~ Mass * Sex + Act + chamber + (batch | date)
expplusglmer <- glmer(CO2.prod.uL.min ~ Sex + Act + chamber + (batch|Age), family= "Gamma", data = mf_exposed)
summary(expplusglmer)

Formula: ulminmg ~ Sex + Act + chamber + (batch | date)

shapiro.test(resid(expplusglmer))
```
##The issue with glmer is that in the paper, there was a designated variable "date" and in the data on dryad, the variable is age, which has a different numerical value (3) as opposed to date (2). This has affected my results and I have not been able to replicate them exactly. Collin describes this as an "oversight" but says that age should work the same. But it doesn't. It's close! But it's not exactly the same. So that's what we're working with right now.

#This one was the closest to the results in the paper: "sex was not a significant predictor of MR (Wald t = 0.24, P = 0.81)
```{r}
glmertry <- glmer(CO2.prod.uL.min ~ Mass * Sex + Act + chamber + (batch|Age), family= "Gamma", data = mf_exposed)

summary(glmertry)
Ulmin ~ Mass * Sex + Act + chamber + (batch | date)
```

#Using chamber as a predictor, no sex, no mass. "chamber was a significant predictor in the exposed condition model of MR (Wald t = 5.84, P<0.0001)." Also shows that activity is significant if in the paper they are using the absolute value of the t-statistic: "activity was a significant predictor of MR (Wald t = 4.47, P<0.0001)."
```{r}
glmercham <- glmer(CO2.prod.uL.min * 100 ~ Act + chamber + (batch|Age), family= "Gamma", data = mf_exposed)
summary(glmercham)

Formula: Ulmin * 100 ~ Act + chamber + (batch | date)
```
#Also close, only if we're using the absolute value of the wald t statistic. "In models with mass as a predictor, but not sex, mass significantly predicted MR in the exposed condition (Wald t = 2.27, P = 0.023)."
```{r}
expmassglmer <- glmer(CO2.prod.uL.min ~ Mass + Act + chamber +(batch|Age), family= "Gamma", data = mf_exposed)
summary(expmassglmer)
```




##Third statistical test: Mixed effects modeling to control for sexual dimorphism. Females are typically heavier.
###Our flies are sexually dimorphic (females are typically heavier). That motivated me to run two different mixed efects models (MEM)
```{r}
massEEErestMF <- lmer(CO2.prod.uL.min ~ Age + Mass + chamber + (1|batch), data= mf_rest)
summary(massEEErestMF)
anova(massEEErestMF, test = 'Chisq')
```
Linear mixed model fit by REML. t-tests use Satterthwaite's method ['lmerModLmerTest']
Formula: CO2.prod.uL.min ~ Age + Mass + chamber + (1 | batch)
   Data: MFRestdata

REML criterion at convergence: -339.8

Scaled residuals:
     Min       1Q   Median       3Q      Max
-2.52532 -0.64051  0.02341  0.57266  2.58492

Random effects:
 Groups   Name        Variance  Std.Dev.
 batch    (Intercept) 1.057e-05 0.003251
 Residual             2.860e-04 0.016911
Number of obs: 72, groups:  batch, 6

Fixed effects:
              Estimate Std. Error         df t value Pr(>|t|)    
(Intercept)  0.0255411  0.0167658 28.6698918   1.523  0.13861    
Age          0.0022451  0.0006146  4.5926158   3.653  0.01707 *  
Mass         0.0052076  0.0047864 67.9091407   1.088  0.28044    
chamber     -0.0020577  0.0005790 63.9814777  -3.554  0.00072 ***

#Based on sex
```{r}
sexEEErestMF <- lmer(CO2.prod.uL.min ~ Age + Sex + chamber + (1|batch), data= mf_rest)
summary(sexEEErestMF)
anova(sexEEErestMF, test = 'Chisq')
```
Linear mixed model fit by REML. t-tests use Satterthwaite's method ['lmerModLmerTest']
Formula: CO2.prod.uL.min ~ Age + Sex + chamber + (1 | batch)
   Data: MFRestdata

REML criterion at convergence: -344.1

Scaled residuals:
     Min       1Q   Median       3Q      Max
-2.35390 -0.62481 -0.01316  0.61877  2.90688

Random effects:
 Groups   Name        Variance  Std.Dev.
 batch    (Intercept) 1.765e-05 0.004201
 Residual             2.635e-04 0.016234
Number of obs: 72, groups:  batch, 6

Fixed effects:
              Estimate Std. Error         df t value Pr(>|t|)    
(Intercept)  0.0450120  0.0106003  5.5007490   4.246 0.006566 **
Age          0.0020453  0.0006294  3.9999995   3.250 0.031387 *  
SexM        -0.0094350  0.0038263 64.0000007  -2.466 0.016360 *  
chamber     -0.0020097  0.0005542 64.0000007  -3.626 0.000571 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Correlation of Fixed Effects:
        (Intr) Age    SexM  
Age     -0.891              
SexM    -0.180  0.000      
chamber -0.340  0.000  0.000
###

#Recreating the figure
#CO2 male and female
```{r}
mf_rest1 <- read.csv("mf_CO2_2.csv", header = TRUE, stringsAsFactors = FALSE)

#define variables
treatment=rep(c('Exposed', 'Unexposed'), 'Exposed'=56, 'Unexposed'=72)
Sex=rep(c('Male', 'Female'), each=64)
CO2=mf_rest1$CO2

data=data.frame(treatment, Sex, CO2)
summary(data)

ggplot(data, aes(x=treatment, y=CO2, fill=Sex)) + 
  geom_boxplot()
```

#Activity male and female
```{r}
mf_rest_act <- read.csv("mf_act_2.csv", header = TRUE, stringsAsFactors = FALSE)

#define variables
treatment=rep(c('Exposed', 'Unexposed'), 'Exposed'=56, 'Unexposed'=72)
Sex=rep(c('Male', 'Female'), each=64)
Activity=mf_rest_act$Activity

data=data.frame(treatment, Sex, Activity)
summary(data)

ggplot(data, aes(x=treatment, y=Activity, fill=Sex)) + 
  geom_boxplot()

```
#CO2 virgin and mated. Differing number of rows here which gave me trouble, so I got rid of one of the exposed virgins with an activity level "NA". This did not work.
```{r}
fem_CO2 <- read.csv("fem_CO2_2.csv", header = TRUE, stringsAsFactors = FALSE)

#define variables
treatment=rep(c('Exposed', 'Unexposed'), 'Exposed'=69, 'Unexposed'=55)
Virginity_status=rep(c('Virgin', 'Mated'), each=62)
CO2=fem_CO2$CO2

data=data.frame(treatment, Virginity_status, CO2)
summary(data)

ggplot(data, aes(x=treatment, y=CO2, fill=Virginity_status)) + 
  geom_boxplot()
```

#Recreating this figure: (insert link to picture) this did not work
```{r}
fem_exposed_mr <- fem_exposed$CO2.prod.uL.min
fem_rest_mr <- fem_rest$CO2.prod.uL.min
male_exposed_mr <- male_exposed$CO2.prod.uL.min
male_rest_mr <- male_rest$CO2.prod.uL.min

fem_exposed_mr <- read.csv("fem_exposed_mr.csv", header = TRUE, stringsAsFactors = FALSE)
fem_rest_mr <- read.csv("fem_rest_mr.csv", header = TRUE, stringsAsFactors = FALSE)
male_exposed_mr <- read.csv("male_exposed_mr.csv", header = TRUE, stringsAsFactors = FALSE)
male_rest_mr <- read.csv("male_rest_mr.csv", header = TRUE, stringsAsFactors = FALSE)

boxplot(fem_exposed_mr, fem_rest_mr, male_exposed_mr, male_rest_mr)

fem_exposed_mr$CO2.prod.uL.min <- as.factor(fem_exposed_mr$CO2.prod.uL.min )
head(fem_exposed_mr)

fem_rest_mr$CO2.prod.uL.min <- as.factor(fem_rest_mr$CO2.prod.uL.min )
head(fem_rest_mr)

male_exposed_mr$CO2.prod.uL.min <- as.factor(male_exposed_mr$CO2.prod.uL.min )
head(male_exposed_mr)

male_rest_mr$CO2.prod.uL.min <- as.factor(male_rest_mr$CO2.prod.uL.min )
head(male_rest_mr)
  
library(ggplot2)
ggplot() + 
  geom_boxplot((data = fem_exposed_mr + fem_rest_mr + male_exposed_mr + male_rest_mr), mapping = aes(col_name, value, fill=index))  + 
  theme( axis.text.x = element_blank()) +
  geom_segment(data=hline1, mapping=aes(x=1, y=-0.5, xend=3, yend=-0.5), size=2) +
  annotate("text", x = 2, y = -1, label = "A") + 
  geom_segment(data=hline1, mapping=aes(x=4, y=-0.5, xend=6, yend=-0.5), size=2) +
  annotate("text", x = 5, y = -1, label = "B")   


cols <- rainbow(3, s = 0.5)
boxplot(x ~ z + y, data = DF2,
        at = c(1:3, 5:7), col = cols,
        names = c("", "A", "", "", "B", ""), xaxs = FALSE)
legend("topleft", fill = cols, legend = c(1,2,3), horiz = T)
```





###Stuff that didn't work...

#Calculating average weight for male flies at rest:
```{r}
male_rest <- mf_rest %>% filter(Sex == "M")
male_rest
summary(male_rest)
```
#Mean mass for males at rest is 2.208, which is closer, but still not the same. I will now try combining all males, both exposed and at rest, to see if I can get the same mean from the study. Since they are in two different .csv files, I will try and combine them into one column using the merge function.

#The merge function did not work for me, so I instead created a new .csv file by copying and pasting values from excel to create a combined file of males at rest and exposed:
```{r}
male_combined <- read.csv("combined_mf.csv", header = TRUE, stringsAsFactors = FALSE)
summary(male_combined$Mass)
```
#Mean mass of all males is 2.093, not 2.35